---
title: "Analyzing larval environmental data"
author: "Ariana S Huffmyer"
date: "11/15/2022"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

This script reads and plots environmental data from Hobo pendant loggers and daily measurements.  

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(seacarb)
library(broom)
library(cowplot)
```

# Daily measurements  

## Read in file  

Read in daily measurements file.  

```{r}
daily<-read_csv("data/environmental/daily_measurements/daily_measurements.csv")
daily$date<-as.Date(daily$date, format="%m/%d/%y")
daily$tank<-as.factor(daily$tank)
daily$tris.date<-as.character(daily$tris.date)
```

## Correct salinity values  

The salinity probe was calibrated on 20221026 by HP to 1000 uS/cm standard. The measurements before this date need to be corrected by applying a multiplication factor to all measurements before 20221026.  

```{r}
#calculate the average salinity values before and after the calibration 
before<-daily%>%
  filter(date<"2022-10-26")%>%
  summarise(correction=mean(sal.psu)); before 
#before = 38.2

after<-daily%>%
  filter(date>"2022-10-25")%>%
  summarise(correction=mean(sal.psu)); after 
#after=35.1

correction<-before-after; correction
#correction of 3.12 psu
```

Apply the correction by subtracting the value from all measurements before 2022-10-26. 
```{r}
daily<-daily%>%
  mutate(sal.psu.cor=ifelse(date<"2022-10-26", sal.psu-correction$correction, sal.psu))
```

We will now use sal.psu.cor for future analyses.  

## Calculate pH  

Calculate the calibration curve from the Tris calibration and calculate pH on the total scale from pH.mV.   
```{r}
pHcalib<-read_csv("data/environmental/daily_measurements/20220924_Tris_Calibration.csv")
pHcalib$tris.date<-as.character(pHcalib$tris.date)

pHSlope<-pHcalib %>%
  nest_by(tris.date)%>%
  mutate(fitpH = list(lm(mVTris~TTris, data = pHcalib))) %>% # linear regression of mV and temp of the tris
  summarise(broom::tidy(fitpH)) %>% # make the output tidy
  select(tris.date, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%# put slope and intercept in their own column
  left_join(daily, ., by="tris.date") %>% # join with the pH sample data
  mutate(mVTris = temp.C*TTris + `(Intercept)`) %>%# calculate the mV of the tris at temperature in which the pH of the tanks were measured
  mutate(pH = pH(Ex=pH.mV,Etris=mVTris,S=sal.psu.cor,T=temp.C)) # calculate pH of the tanks using the pH seacarb function

```

pH is now calculated as Total pH in the "pH" column. Now select the desired columns to analyze further.  

```{r}
daily_calc<-pHSlope%>%
  select(date, time, tank, temp.C, pH, sal.psu.cor, flow.mL.5s, par)
```

## Calculate flow  

Calculate flow to total mL per minute rather than 5 sec.  
```{r}
daily_calc<-daily_calc%>%
  mutate(flow.L.min=(flow.mL.5s*12)/1000)%>%
  select(!flow.mL.5s)
```

## Change to long format

Change data format to long format 
```{r}
daily_calc.long <-daily_calc %>% pivot_longer(cols=temp.C:flow.L.min,
  names_to = "metric",
  values_to = "value")
```

Add in treatment information.  
```{r}
daily_calc.long<-daily_calc.long%>%
  mutate(treatment=ifelse(grepl('^A', tank), 'High', 'Ambient'))%>%
  mutate(treatment=ifelse(grepl('^P', tank), 'Parent', treatment))
```

## Plot metrics of interest  

Plot by tank colored by temperature treatment.      
```{r}
daily_tanks<-daily_calc.long %>%
  ggplot(aes(x=date, y=value, colour=treatment))+
  geom_point(size=2)+
  geom_vline(xintercept = as.numeric(ymd("2022-10-19")), linetype="dashed", 
                color = "black", size=0.5)+
  geom_vline(xintercept = as.numeric(ymd("2022-10-24")), linetype="solid", 
                color = "black", size=0.5)+
  scale_colour_manual(values=c("blue", "red", "gray"))+
  xlab("Date")+
  facet_grid(metric ~ ., scales = "free")+
  scale_fill_brewer(palette = "Dark2")+
  theme_bw(); daily_tanks
  
ggsave(filename="figures/environmental/tank_daily_measurements.png", plot=daily_tanks, dpi=300, width=6, height=8, units="in")

```

Plot by treatment summarizing tanks within each treatment.    
```{r}
daily_treatment<-daily_calc.long %>%
  ggplot(aes(x=date, y=value, colour=treatment, group=interaction(treatment, date)))+
  geom_point()+
  geom_boxplot()+
  geom_vline(xintercept = as.numeric(ymd("2022-10-19")), linetype="dashed", 
                color = "black", size=0.5)+
  geom_vline(xintercept = as.numeric(ymd("2022-10-24")), linetype="solid", 
                color = "black", size=0.5)+
  scale_colour_manual(values=c("blue", "red", "gray"))+
  xlab("Date")+
  facet_grid(metric ~ ., scales = "free")+
  scale_fill_brewer(palette = "Dark2")+
  theme_bw(); daily_treatment

ggsave(filename="figures/environmental/treatment_daily_measurements.png", plot=daily_treatment, dpi=300, width=6, height=8, units="in")
```

## Test for differences between tanks  

Obtain a short format dataset.  
```{r}
daily_calc.short<-daily_calc.long%>%
  spread(metric, value)
```

Test for differences between metrics in High tanks during the exposure period.  
```{r}
daily_calc.short%>%
  filter(treatment=="High")%>%
  filter(date>"2022-10-23")%>%
  aov(temp.C~tank, data=.)%>%
  summary()
#no difference 

daily_calc.short%>%
  filter(treatment=="High")%>%
  filter(date>"2022-10-23")%>%
  aov(pH~tank, data=.)%>%
  summary()
#no difference 

daily_calc.short%>%
  filter(treatment=="High")%>%
  filter(date>"2022-10-23")%>%
  aov(sal.psu.cor~tank, data=.)%>%
  summary()
#no difference 

daily_calc.short%>%
  filter(treatment=="High")%>%
  filter(date>"2022-10-23")%>%
  aov(par~tank, data=.)%>%
  summary()
#no difference

daily_calc.short%>%
  filter(treatment=="High")%>%
  filter(date>"2022-10-23")%>%
  aov(flow.L.min~tank, data=.)%>%
  summary()
#no difference 
```

There is no difference between tanks within the high temperature treatment.  

Test for differences in ambient tanks during the exposure period.  
```{r}
daily_calc.short%>%
  filter(treatment=="Ambient")%>%
  filter(date>"2022-10-23")%>%
  aov(temp.C~tank, data=.)%>%
  summary()
#no difference 

daily_calc.short%>%
  filter(treatment=="Ambient")%>%
  filter(date>"2022-10-23")%>%
  aov(pH~tank, data=.)%>%
  summary()
#no difference 

daily_calc.short%>%
  filter(treatment=="Ambient")%>%
  filter(date>"2022-10-23")%>%
  aov(sal.psu.cor~tank, data=.)%>%
  summary()
#no difference 

daily_calc.short%>%
  filter(treatment=="Ambient")%>%
  filter(date>"2022-10-23")%>%
  aov(par~tank, data=.)%>%
  summary()
#no difference

daily_calc.short%>%
  filter(treatment=="Ambient")%>%
  filter(date>"2022-10-23")%>%
  aov(flow.L.min~tank, data=.)%>%
  summary()
#no difference 
```

## Test for differences between treatments  
```{r}
daily_calc.short%>%
  filter(!treatment=="Parent")%>%
  filter(date>"2022-10-23")%>%
  aov(temp.C~treatment, data=.)%>%
  summary()
#temperature is different between treatments

daily_calc.short%>%
  filter(!treatment=="Parent")%>%
  filter(date>"2022-10-23")%>%
  aov(pH~treatment, data=.)%>%
  summary()
#pH is different between treatments

daily_calc.short%>%
  filter(!treatment=="Parent")%>%
  filter(date>"2022-10-23")%>%
  aov(flow.L.min~treatment, data=.)%>%
  summary()
#flow is not different

daily_calc.short%>%
  filter(!treatment=="Parent")%>%
  filter(date>"2022-10-23")%>%
  aov(par~treatment, data=.)%>%
  summary()
#light is not different

daily_calc.short%>%
  filter(!treatment=="Parent")%>%
  filter(date>"2022-10-23")%>%
  aov(sal.psu.cor~treatment, data=.)%>%
  summary()
#salinity is not different 
```

View mean difference in pH between treatments.  
```{r}
daily_calc.short%>%
  filter(date>"2022-10-23")%>%
  group_by(treatment)%>%
  summarise(mean=mean(pH, na.rm=TRUE))
```
pH is 0.04 units lower in the high temperature treatment than the ambient treatment.  

View mean difference in temperature between treatments.  
```{r}
daily_calc.short%>%
  filter(date>"2022-10-23")%>%
  group_by(treatment)%>%
  summarise(mean=mean(temp.C, na.rm=TRUE))
```
Temperature is 2.4° higher in the high temperature treatment than the ambient temperature values, which was the target difference.    

## Summarize daily measurements  

Add a time period for spawning, embryo rearing, and larval exposure.  

```{r}
daily_calc.short<-daily_calc.short%>%
  mutate(period=ifelse(date>"2022-10-23", 'Exposure', 'Embryo Rearing'))%>%
  mutate(period=ifelse(date<"2022-10-19", 'Spawning', period))

daily_calc.short$period<-factor(daily_calc.short$period, levels=c("Spawning", "Embryo Rearing", "Exposure"))
```

Calculate descriptive statistics   
```{r}
daily_calc.short%>%
  group_by(period, treatment)%>%
  select(!tank)%>%
  select(!date)%>%
  select(!time)%>%
  summarise(across(everything(), list(mean = mean, sd = sd), na.rm = TRUE))

```

# Logger analysis    

## Read in files  

Read in Hobo Pendant files  
```{r}
pendant.files <- list.files(path="data/environmental/loggers/", pattern="*.xlsx", full.names=TRUE)

pendants <- setNames(pendant.files, pendant.files) %>%
   map_dfr(read_excel, .id = "logger") %>%
   rename(DateTime=`Date-Time (French Polynesia Standard Time)`, temp.C=`Ch: 1 - Temperature   (°C )`, light.lux=`Ch: 2 - Light   (lux)`)%>%
  select(!`#`)%>%
  mutate(logger=sub(".*/", "", logger))%>% #remove the path name
  mutate(logger=sub("_.*", "", logger)) #keep only the serial number 

pendants$DateTime<-as.POSIXct(pendants$DateTime, format="%m/%d/%y %H:%M:%S", tz="Pacific/Tahiti")

#adjust time to correct values and force time zone
pendants$DateTime <- force_tz(pendants$DateTime, tzone = "Pacific/Tahiti")
```

List the logger serial numbers in this dataframe.  
```{r}
levels(as.factor(pendants$logger))
```
21002974
21002979
21002981
21002982
21335980
21335987

## Read in treatment information  

Read in metadata that assigns a pendant serial number to a tank number and/or treatment.  
```{r}
metadata<-read_csv("data/environmental/loggers/logger_metadata_larvae.csv")
metadata$logger<-as.character(metadata$logger)
```

Assign treatment information to logger data.  
```{r}
pendants<-left_join(pendants, metadata, by="logger")
pendants$tank<-as.factor(pendants$tank)
```

## Apply tempreature calibrations  

Read in calibration output file with the intercept and coefficients for calibration. This calibration file is generated by the E5 project here https://github.com/urol-e5/apulchra_metabolism. The output files can be found here https://github.com/urol-e5/apulchra_metabolism/tree/main/output/environmental.  
```{r}
temp_cal<-read_csv("https://raw.githubusercontent.com/urol-e5/apulchra_metabolism/main/output/environmental/temperature_logger_calibrations.csv")
temp_cal$logger<-as.character(temp_cal$logger)
```

Generate a calibrated temperature column in our logger dataframe.  
```{r}
pendants<-left_join(pendants, temp_cal, by="logger")

pendants<-pendants%>%
  mutate(temp.calibrated=temp.C*coef+intercept)%>%
  select(!coef)%>%
  select(!intercept)
```

## Apply light calibrations  

Read in calibration output file with the intercept and coefficients for calibration that was calculated as an offset to the mean Lux value from all pendant loggers.    
```{r}
light_cal<-read_csv("https://raw.githubusercontent.com/urol-e5/apulchra_metabolism/main/output/environmental/light_logger_calibrations.csv")
light_cal$logger<-as.character(light_cal$logger)
```

Apply calibration values.  
```{r}
pendants<-left_join(pendants, light_cal, by="logger")

pendants<-pendants%>%
  mutate(lux.calibrated=light.lux*coef+intercept)%>%
  select(!coef)%>%
  select(!intercept)
```

Apply a lux to PPFD (µmol m-2 s-1) conversion factor (0.0185) as provided by Apogee for sunlight. https://www.apogeeinstruments.com/conversion-ppfd-to-lux/ 
```{r}
pendants$par<-pendants$lux.calibrated*0.0185
```

## Plot temperature    

Plot raw data over time by tank colored by treatment.   
```{r}
temp_plot<-pendants%>%
  
  ggplot(aes(x=DateTime, y=temp.calibrated, colour=tank))+
  geom_line()+
  scale_colour_manual(values=c("red", "red", "red", "blue", "blue", "blue"))+
    geom_vline(xintercept = as.POSIXct("2022-10-19"), linetype="dashed", 
                color = "black", size=0.5)+
  geom_vline(xintercept = as.POSIXct("2022-10-24"), linetype="solid", 
                color = "black", size=0.5)+
  theme_classic(); temp_plot
```

Remove temperature data from logger read outs and tank cleaning.  



ADD PLOTS FOR DAILY FLUCTUATIONS 
REMOVE LOW VALUES DURING READ OUT 







## Plot light     

Plot over time by tank in par units  
```{r}
light_plot<-pendants%>%
  ggplot(aes(x=DateTime, y=par, colour=tank))+
  ylab("PAR (umol m2 s)")+
  scale_colour_manual(values=c("red", "red", "red", "blue", "blue", "blue"))+
  geom_vline(xintercept = as.POSIXct("2022-10-19"), linetype="dashed", 
                color = "black", size=0.5)+
  geom_vline(xintercept = as.POSIXct("2022-10-24"), linetype="solid", 
                color = "black", size=0.5)+
  geom_line()+
  theme_classic(); light_plot
```

Max daily PAR was ~150-250 in larval tanks.  

Plot by lux
```{r}
lux_plot<-pendants%>%
  ggplot(aes(x=DateTime, y=lux.calibrated, colour=tank))+
  ylab("Light (lux)")+
    scale_colour_manual(values=c("red", "red", "red", "blue", "blue", "blue"))+
    geom_vline(xintercept = as.numeric(as.POSIXct("2022-10-19")), linetype="dashed", 
                color = "black", size=0.5)+
  geom_vline(xintercept = as.numeric(as.POSIXct("2022-10-24")), linetype="solid", 
                color = "black", size=0.5)+
  geom_line()+
  theme_classic(); lux_plot
```


ADD CODE TO SAVE FILES OF INTEREST 
ADD LIGHT OVER COURSE OF DAILY FLUCTUATION 