---
title: Respirometry Plotting and Analysis  
author: "AS Huffmyer"
date: '2022'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
This script analyzes and plots data for Acropora pulchra larval metabolism 2022 respirometry data. 

# **Setup**  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('effects')
library('emmeans')
library('multcomp')
```

# **Data visualization and manipulation**  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
PRdata<-read.csv("output/respirometry/oxygen_P_R_calc.csv") #load data
```

Separate project specific data.  
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
PRdata<-PRdata[!is.na(PRdata$Type),]

#format columns
PRdata$Meas.Temp<-as.factor(PRdata$Meas.Temp)
PRdata$Tank<-as.factor(PRdata$Tank)
PRdata$Treatment.Temp<-as.factor(PRdata$Treatment.Temp)
PRdata$Symbiont<-as.factor(PRdata$Symbiont)
PRdata$Treatment.Code<-as.factor(PRdata$Treatment.Code)
PRdata$Days.Exposure<-as.factor(PRdata$Days.Exposure)
PRdata$SDR<-as.factor(PRdata$SDR)
PRdata$Run<-as.factor(PRdata$Run)
```

Calculate a P:R ratio using gross photosynthesis.  
```{r, warning=FALSE, message=FALSE}
PRdata$ratio<-abs(PRdata$GP.nmol.org.min)/abs(PRdata$R.nmol.org.min) #calculate ratio with absolute values
#remove outliers detected by values of P:R ratio data
boxplot(PRdata$ratio)
PRdata<-PRdata%>%filter(ratio < 5)
boxplot(PRdata$ratio)
```

Look for outliers in the data.  

```{r}
boxplot(PRdata$R.nmol.org.min)
#PRdata<-PRdata%>%filter(R.nmol.org.min < 10)
#boxplot(PRdata$R.nmol.org.min)

boxplot(PRdata$P.nmol.org.min)
#PRdata<-PRdata%>%filter(P.nmol.org.min < 10)
#boxplot(PRdata$P.nmol.org.min)

boxplot(PRdata$GP.nmol.org.min)
#PRdata<-PRdata%>%filter(GP.nmol.org.min < 10)
#boxplot(PRdata$GP.nmol.org.min)
```

No apparent outliers in the data.  

Remove runs 9, 10, 11, 12. 

```{r}
#remove<-c("9", "10", "11", "12")
#PRdata<-PRdata%>%
  #filter(!Run %in% remove)
```

# **Plotting**  

## Respiration  

Plot data by Run   
```{r}
r_plot<-PRdata %>%
    group_by(Meas.Temp, Treatment.Temp, Run)%>%
    dplyr::summarise(mean=mean(abs(R.nmol.org.min)), sd=sd(abs(R.nmol.org.min)), N=length(R.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(Run), y = abs(mean), group=interaction(Meas.Temp, Treatment.Temp))) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(fill=Meas.Temp, group=interaction(Treatment.Temp, Meas.Temp), shape=Treatment.Temp, colour=Meas.Temp), size=6, position = position_dodge(0.2)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Treatment.Temp, Meas.Temp)), width=0, linetype="solid", position=position_dodge(0.2), size=1.3, color="black")+
    xlab("Run Number") + 
    #scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    #scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot

#ggsave("Mcap2021/Figures/Respiration/Respiration.png", r_plot, dpi=300, w=5, h=5, units="in")
```

Plot data by Run as boxplot  
```{r}
r_plot<-PRdata %>%
    group_by(Meas.Temp, Treatment.Temp, Run)%>%
    
    ggplot(., aes(x = as.factor(Run), y = abs(R.nmol.org.min), group=interaction(Meas.Temp))) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(group=interaction(Meas.Temp, Run), colour=Meas.Temp), position = position_dodge(0.8)) + 
    geom_point(aes(fill=Meas.Temp, colour=Meas.Temp), size=3, position = position_dodge(0.8)) +
    xlab("Run Number") + 
    #scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    #scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot

#ggsave("Mcap2021/Figures/Respiration/Respiration.png", r_plot, dpi=300, w=5, h=5, units="in")
```

Plot data as an interaction plot   
```{r}
r_plot<-PRdata %>%
    group_by(Meas.Temp, Treatment.Temp, Days.Exposure)%>%
    dplyr::summarise(mean=mean(abs(R.nmol.org.min)), sd=sd(abs(R.nmol.org.min)), N=length(R.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(Days.Exposure), y = abs(mean), group=interaction(Meas.Temp, Treatment.Temp))) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(fill=Meas.Temp, group=interaction(Treatment.Temp, Meas.Temp), shape=Treatment.Temp, colour=Meas.Temp), size=6, position = position_dodge(0.2)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Treatment.Temp, Meas.Temp)), width=0, linetype="solid", position=position_dodge(0.2), size=1.3, color="black")+
    xlab("Days Exposure") + 
    #scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    #scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot

#ggsave("Mcap2021/Figures/Respiration/Respiration.png", r_plot, dpi=300, w=5, h=5, units="in")
```


After run 8 things get weird - higher values and more variable in runs 9-12 and 13-16???


## Photosynthesis (Net)  

Plot data as an interaction plot  
```{r}
p_plot<-PRdata %>%
    group_by(Meas.Temp, Treatment.Temp, Days.Exposure)%>%
    dplyr::summarise(mean=mean(abs(P.nmol.org.min)), sd=sd(abs(P.nmol.org.min)), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(Days.Exposure), y = abs(mean), group=interaction(Meas.Temp, Treatment.Temp))) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(fill=Meas.Temp, group=interaction(Treatment.Temp, Meas.Temp), shape=Treatment.Temp, colour=Meas.Temp), size=6, position = position_dodge(0.2)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Treatment.Temp, Meas.Temp)), width=0, linetype="solid", position=position_dodge(0.2), size=1.3, color="black")+
    xlab("Days Exposure") + 
    #scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    #scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("Net P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); p_plot
```


































## Generate final figure  

```{r}
full_fig<-plot_grid(r_plot, p_plot, gp_plot, pr_plot, ncol=4, nrow=1, rel_heights= c(1,1,1,1), rel_widths = c(1,1,1,1.4), label_y=1, align="h", labels=c("A", "B", "C", "D"), label_size=18)

ggsave(filename="Mcap2021/Figures/Respiration/respirometry_fig.png", plot=full_fig, dpi=500, width=18, height=6, units="in")
```

```{r}
full_fig_box<-plot_grid(r_plot_2, p_plot_2, gp_plot_2, pr_plot_2, ncol=4, nrow=1, rel_heights= c(1,1,1,1), rel_widths = c(1,1,1,1.4), label_y=1, align="h")

ggsave(filename="Mcap2021/Figures/Respiration/respirometry_fig_box.png", plot=full_fig_box, dpi=500, width=18, height=6, units="in")
```

# **Analyze data**    

## Respiration     

Build linear mixed effect model and examine for Respiration.      
```{r, results=TRUE, warning=FALSE, message=FALSE}
Rmodel1<-lmer(R.nmol.org.min~Treatment*Temperature + (1|Run) + (1|Tank), data=PRdata) #run as random
anova(Rmodel1, type="II") 
summary(Rmodel1)
```

Run only as regular ANOVA.  
```{r}
summary(aov(R.nmol.org.min~Treatment*Temperature, data=PRdata))
```
Same results.  

Check assumptions of model for residual normality and variance. Passes assumptions.    

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(Rmodel1))
leveneTest(residuals(Rmodel1)~Treatment*Temperature, data=PRdata)
```

Conduct post hoc test.  

```{r}
emm<-emmeans(Rmodel1, ~Treatment*Temperature)
cld(emm)
```


## Net Photosynthesis   

Build linear mixed effect model and examine for Photosynthesis       
```{r, results=TRUE, warning=FALSE, message=FALSE}
Pmodel1<-lmer(P.nmol.org.min~Treatment*Temperature + (1|Run) + (1|Tank), data=PRdata) #run nested within day
anova(Pmodel1, type="II")
summary(Pmodel1)
```

Run only as regular ANOVA.  
```{r}
summary(aov(P.nmol.org.min~Treatment*Temperature, data=PRdata))
```
In this analysis, treatment is significant.  

Check assumptions of model for residual normality and variance. Violates variance assumption, return to this.    

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(Pmodel1))
leveneTest(residuals(Pmodel1)~Treatment*Temperature, data=PRdata)
```

Conduct post hoc test.  

```{r}
emm<-emmeans(Pmodel1, ~Treatment*Temperature)
cld(emm)
```
 

## Gross Photosynthesis  

Build linear mixed effect model and examine for GP       
```{r, results=TRUE, warning=FALSE, message=FALSE}
GPmodel1<-lmer(GP.nmol.org.min~Treatment*Temperature + (1|Run) + (1|Tank), data=PRdata) #run nested within day
anova(GPmodel1, type="II")
summary(GPmodel1)
```

Run only as regular ANOVA.  
```{r}
summary(aov(GP.nmol.org.min~Treatment*Temperature, data=PRdata))
```
Same results.    

Check assumptions of model for residual normality and variance. Passes.    

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(GPmodel1))
leveneTest(residuals(GPmodel1)~Treatment*Temperature, data=PRdata)
```

Conduct post hoc test.  

```{r}
emm<-emmeans(GPmodel1, ~Treatment*Temperature)
cld(emm)
```

## P:R    
 
Build linear mixed effect model and examine for PR ratio       
```{r, results=TRUE, warning=FALSE, message=FALSE}
PRmodel1<-lmer(ratio~Treatment*Temperature + (1|Run) + (1|Tank), data=PRdata) #run as random
anova(PRmodel1, type="II")
summary(PRmodel1)
```

Run only as regular ANOVA.  
```{r}
summary(aov(ratio~Treatment*Temperature, data=PRdata))
```
Same results.   

Check assumptions of model for residual normality and variance. Meets assumptions.  

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(PRmodel1))
leveneTest(residuals(PRmodel1)~Treatment*Temperature, data=PRdata)
```

Conduct post hoc test.  

```{r}
emm<-emmeans(PRmodel1, ~Treatment*Temperature)
cld(emm)
```

### Generate summary tables of mean and SE for all variables  

Generate summary of all respiration data.  

```{r}
summary<-PRdata%>%
  group_by(Temperature, Treatment)%>%
  summarise(N=length(R.nmol.org.min),
            Mean_R=mean(R.nmol.org.min), 
            SE_R=sd(R.nmol.org.min)/sqrt(length(R.nmol.org.min)),
            Mean_P=mean(P.nmol.org.min), 
            SE_P=sd(P.nmol.org.min)/sqrt(length(P.nmol.org.min)),
            Mean_GP=mean(GP.nmol.org.min), 
            SE_GP=sd(GP.nmol.org.min)/sqrt(length(GP.nmol.org.min)),
            Mean_PR=mean(ratio), 
            SE_PR=sd(ratio)/sqrt(length(ratio)))

summary%>%
  write_csv(., "Mcap2021/Output/Respiration/mean_respiration.csv")
```